---
title: "Survey and Commercial Fishing"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description of data

The survey biomass and catch (landings) of the species modelled in Atlantis are extracted from the svdbs database (using survdat) cfdbs (using comlandr) and assembled as a single data frame in long format

```{r code, echo = T,eval= F}
library(magrittr)
# Connect to the database
channel <- dbutils::connect_to_database("servername","user")
# list species of interest
speciesList <- map_functional_group(channel)
# grab data from databases and assemble
data <- extract_landings_biomass(channel,species = speciesList)
```
```{r data, echo=F, eval = T}
dplyr::as_tibble(readRDS(here::here("data-raw","survComData.RDS")))
```


## Methods

Species assigned to the initial Atlantis functional group designations (from v1.0 - found in `initialFunctionalGroupNames.csv`) were used to look up SVSPP, NESPP3, Species_ITIS codes

The `map_functional_groups.r` function extends these tables (from Atlantis v1.0)
```{r initialcsv, echo=T,eval=T, message=F}
readr::read_csv(here::here("data-raw","initialFunctionalGroupNames.csv"))
readr::read_csv(here::here("data-raw","Atlantis_1_5_groups_svspp_nespp3.csv"))
```
to this table 
```{r finalTable, echo=T,eval=T, message=F}
readr::read_csv(here::here("data-raw","FunctionalGroupNames.csv"))
```

Each species is cross-referenced in the svdbs and cfdbs databases to obtain Species_itis codes and scientific names

The list of species codes is then used to extract commercial landings (`comlandr`) and biomass (`survdat`) data using `extract_landings_biomass.r`

