---
title: "Atlantis Diagnostics Development Documentation"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: modreview.bib
csl: plos.csl
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

Here we document the development Atlantis diagnostics code to determine whether the model is meeting define performance and review criteria. Code we be developed to evaluate the model output against detailed performance criteria developed in @kaplan_guinea_2016:

1. All functional groups persist

1. Model stabilizes for the last ~20 years of an unfished, unperturbed 80-100 year run

1. Hindcast period established where we have survey/assessment time series with error bounds

1. Species groups totaling ~80% of system biomass should qualitatively match hindcast biomass trends

1. Patterns of temporal variability captured (emergent or forced with e.g. recruitment time series)

1. Productivity for groups totaling ~80% of system biomass should qualitatively match FMSY or life history expectations

1. Natural mortality decreases with age for majority of groups

1. Age and length structure qualitatively matches expectations for majority of groups

1. Diet predicted qualitatively matches empirical diet comp for majority of groups

An R function for each criterion is developed below, and a wrapper that runs all functions will be developed and tested here.

We use these R libraries, with non-CRAN package installation instructions in comments. 

```{r dependencies}

#devtools::install_github("noaa-edab/ecodata",build_vignettes=TRUE) 
# suggested to use remotes::install_github instead
#devtools::install_github("r4atlantis/atlantisom")

library(ecodata)
library(tidyverse)
library(atlantisom)
library(here)

```

Configure files to be read in here (can have different files to source)

```{r config}
# set up files to be read in
# the below can go into a config file to be sourced

d.name <- here("diagnostics", "testfiles") #folder with files below; e.g. here("atlantisoutput","currentrun")
functional.groups.file <- "NeusGroups_v15_unix_RM.csv"
biomass.pools.file <- ".nc"
biol.prm.file <- "at_biol_neus_v15_scaled_diet_20190710a_lt.prm"
box.file <- ".bgm"
initial.conditions.file <- "RMinit4_2019.nc"
run.prm.file <- "at_run_neus_v15_RM_scale_0503.xml"
scenario.name <- "OLDNEUS"
bioind.file <- "outputOLDNEUSBiomIndx.txt"
catch.file <- "outputOLDNEUSCatch.txt"

```

Get functional group names for other functions

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```


### Persistence test

Setup but need the real output file because old one doesnt match new group names.

```{r persist, eval=FALSE}

# read the output atlantis biom.txt file

atBtxt <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
groupslookup <- funct.groups %>%
  filter(IsTurnedOn > 0)

atBtxttidy <- atBtxt %>%
  select(Time:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) 

# flag any groups with biomass below a threshold 

```

### Stability test

Model reaches steady state for last ~20 years of an unperturbed, unfished ~100 year run.

```{r stability}

# assumes biom.txt output already read in

# look for non-significant linear slope over last 20 years? 30 years would be better

```

### Establish hindcast period

See discussion [here](https://github.com/NOAA-EDAB/atneus_RM/issues/7) and [here](https://github.com/NOAA-EDAB/atneus_RM/wiki/Meeting-minutes#20191002).

The period is defined as 1980-2010 for trend comparison. Change it here to see different results:

```{r hindcast-def}

hindcast <- c(1980:2010)

```

Data sources include ecodata and stock assessment outputs for major species.

### Interim step: which species to include and what data are in the reference set for comparison

Not all species need to match, we will first determine the most important subset(s). Based on discussion, we want to track the species the comprise 80% of biomass (and 80% of catch and 80% of revenue). 

Code to evaluate which species are most important based on biomass, catch, and revenue, union and intersection.

```{r which-species}

# first past B, catch, revenue, use ecodata

survbio <- ecodata::nefsc_survey_disaggregated %>%
  filter(Time %in% hindcast) %>%
  group_by(comname) %>%
  summarize(avgkgtow = mean(kg.per.tow, na.rm = T)) %>%
  mutate(prop = avgkgtow/sum(avgkgtow, na.rm = T)) %>%
  arrange(desc(prop))

hibiosp <- survbio$comname[cumsum(survbio$prop)<0.81 & !is.na(cumsum(survbio$prop))]

# comdat and bennet are already aggregated in ecodata

# need comland, should be able to run similar code as above to get catch and revenue

# add recreational catch and rec value/n participants?

```

Species comprising 80% of survey kg per tow averaged over `r hindcast[1]` to `r hindcast[length(hindcast)]`:

`r hibiosp`


Code to develop the hindcast comparison (call it "reference") dataset:

### Species qualitatively match trends

Key species to match and their most representative trend datasets are determined in analyses above, and the code that looks for the match between this reference set and Atlantis output is developed here. By match we mean significance and direction only. 

```{r trend-match}

# assumes biom.txt output already read in



```



## References